--
title: "Myakka extrapolation"
author: "Sophie McKee"
date: "2024-07-01"
output: word_document
---

```{r setup, include=FALSE}
# Clean existing variables
rm(list = ls(all.names = TRUE))
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

```

## Myakka - data upload

```{r upload data, warning=FALSE}
library(readxl)
setwd("~/Library/CloudStorage/OneDrive-USDA/Documents/USACE FY24 - 25/Myakka/DataAnalysis")
# input damage data
data <- read_excel("MRSP_Tidy_Field_RS_Master.xlsx") %>%
    mutate(Severity = case_when(Severity == 'na' ~ NA,
                           TRUE ~ Severity)) %>% 
    mutate(Age = case_when(Age == 'na' ~ NA,
                           TRUE ~ Age)) %>%
    mutate(Wetland_ATTRIBUTE = case_when(Wetland_ATTRIBUTE == 'na' ~ NA,
                           TRUE ~ Wetland_ATTRIBUTE)) %>%
    mutate(WETLAND_TY = case_when(WETLAND_TY == 'na' ~ NA,
                           TRUE ~ WETLAND_TY)) %>%
    mutate(Wetland_ACRES = case_when(Wetland_ACRES == 'na' ~ NA,
                           TRUE ~ Wetland_ACRES)) %>% 
    mutate(Community_ACREAGE = case_when(Community_ACREAGE == 'na' ~ NA,
                           TRUE ~ Community_ACREAGE)) %>% 
    mutate(EC_TYPE = case_when(EC_TYPE == 'na' ~ NA,
                           TRUE ~ EC_TYPE)) 

table(data$WETLAND_TY)
dataAll <- data
# Select only wetlands
dataW <- dataAll[!is.na(dataAll$Wetland_ATTRIBUTE),]

```


## A. Damage extrapolation circumscribed to "before"


```{r }
data.B <- dataAll[dataAll$Before_After_Control == "Before",]

```

```{r}
# Create a dataframe "all_quadrats" recording the center of all quadrats in the transects, along with their 
# characteristics (number of cells and successes, distance to road and water, forest type)
all_quadrats <- data.B %>% 
  group_by(Transect, Transect_Length_in_m, Before_After_Control, Observer) %>% 
  # Aggregate to the left and right (Width)
  summarise(successes = sum(Damage_Coverage), water = sum(Water_Coverage), total = n(), 
             dist_FIRELINE_m = mean(NEAR_DIST_Fireline_meters), 
             dist_AG_m = mean(NEAR_DIST_2021_NLCD_AG_meters), 
             NDWI = mean(NDWI), 
             dist_Burned_Unit_m = mean(NEAR_DIST_Burned_Unit_meters)) %>% 
  # Aggregateby series of 4 m (Length)
  group_by(Transect) %>% 
  mutate(quadrat = (row_number() - 1) %/% 4) %>%     
  group_by(Transect, Before_After_Control, quadrat) %>% 
  summarise(successes = sum(successes), water = sum(water), total = sum(total), Transect_Length_in_m = min(Transect_Length_in_m), 
             dist_FIRELINE_m = mean(dist_FIRELINE_m), 
             dist_AG_m = mean(dist_AG_m), 
             NDWI = mean(NDWI), 
             dist_Burned_Unit_m = mean(dist_Burned_Unit_m)) %>% 
  # Keep only complete quadrats (4m X 4m)
  filter(total == 32) %>% 
  # Keep only quadrats that are not covered 100% in water
  filter(water != 32) %>% 
  mutate(cells = total - water, failures = cells - successes) %>% 
  select(-total, -water, -cells)

# Width invariant data
Transect_data <- data.B[data.B$Transect_Width_in_m == 1 & data.B$Transect_Left_Right == "Right", c("Transect", "Before_After_Control", "Transect_Length_in_m", "Observer", "Treatment_Non_treatment", "Wetland_ATTRIBUTE", "WETLAND_TY", "EC_TYPE", "Park_Sample_Location")]

all_quadrats <- left_join(all_quadrats, Transect_data, by = c("Transect", "Transect_Length_in_m"))

all_quadrats$rate <- all_quadrats$successes / (all_quadrats$successes + all_quadrats$failures)

y_all <- cbind(all_quadrats$successes, all_quadrats$failures)

write_csv(all_quadrats, "all_quadrats.csv")
```
Plotting success rate as a function of distance to water and road

```{r}

ggplot(all_quadrats) +
    geom_point(aes(x = dist_Burned_Unit_m, y = rate)) +
    theme_bw() +
    labs(
          x = "Distance to fire (m)",
          y = "Damage ratio"
      )

ggplot(all_quadrats) +
    geom_point(aes(x = dist_FIRELINE_m, y = rate)) +
    theme_bw() +
    labs(
          x = "Distance to fireline (m)",
          y = "Damage ratio"
      )

ggplot(all_quadrats) +
    geom_point(aes(x = dist_AG_m, y = rate)) +
    theme_bw() +
    labs(
          x = "Distance to ag (m)",
          y = "Damage ratio"
      )
```



```{r}

# Fit a GAM model
library(lme4)

all_quadrats$dist_AG_100 <- all_quadrats$dist_AG_m / 100
all_quadrats$dist_FIRELINE_100 <- all_quadrats$dist_FIRELINE_m / 100
all_quadrats$dist_Burned_Unit_100 <- all_quadrats$dist_Burned_Unit_m / 100


m1 <- glmer(y_all ~ dist_Burned_Unit_100 + dist_FIRELINE_100 + dist_AG_100 + Park_Sample_Location + EC_TYPE + (1|Transect) + (1|Observer), family = binomial, data = all_quadrats)
summary(m1)

summary(all_quadrats$dist_water100)
summary(sample_quadrats$dist_water100)
anova(m2, m3, text = "Chisq")

library(sjPlot)
tab_model(m3)


```


```{r}
##########
## Other options to be investigated next year?
##########

library(caret)

#specify the cross-validation method
ctrl <- trainControl(method = "LOOCV")

#fit a regression model and use LOOCV to evaluate performance
model <- train(y ~ x1 + x2, data = df, method = "lm", trControl = ctrl)

#view summary of LOOCV               
print(model)



# STAN 
library(rstanarm)
bglm_1 <- stan_glm(y ~ water100 + treatment,
                              data = sample_quadrats,
                              family = binomial(link = "logit"), 
                              prior = student_t(df = 7, 0, 5), 
                              prior_intercept = student_t(df = 7, 0, 5),
                              cores = 2, seed = 12345)
summary(bglm_1)

(bglm_2 <- update(bglm_1, formula. = . ~ . + DUST))

summary(bglm_2)


# Vegetation type, all roads, thickets, soils



loo_bglm_1 <- loo(bglm_1)
loo_bglm_2 <- loo(bglm_2)

comp <- loo_compare(loo_bglm_1, loo_bglm_2)
print(comp, digits = 2)
```
# Extrapolation to 2 plots
```{r}
# Upload grid data
grid <- read_excel("transect_and_grid_data.xlsx", 
    sheet = "4x4_grid_points", col_types = c("skip", 
        "numeric", "numeric", "text", "text"))

grid <- grid %>% rename(dist_to_road = distance_to_road_meters)
grid <- grid %>% rename(dist_to_water = distance_to_water_meters)
grid <- grid %>% rename(treatment = management)
grid$dist_water100 <- grid$dist_to_water/100
grid$dist_road100 <- grid$dist_to_road/100
grid$mast <- ifelse(grid$`forest_type` %in% c("Overcup Oak/Bitter Pecan",
                                                                    "Overcup Oak/Bitter Pecan & Nuttall Oak/American Elm/Green Ash/ Sugarberry",
                                                                    "Overcup Oak/Bitter Pecan/Sugarberry"), 1, 0)

# Model to be used
summary(m3)

# Predict to the 2 plots
# This method works to estimate the mean of the damage ratio
# pred <- predict(m3, newdata=grid, type = "response",re.form=NA)  # list with two elements fit and se.fit
# pred <- predict(m3, newdata=gridT, type = "response")  # list with two elements fit and se.fit

# We randomly assign a transect number to each quadrat because that's the only way to have confidence intervals
transect_pool_nontreat <- LETTERS[1:5]
transect_pool_treat <- LETTERS[6:10]

pool_nontreat <- sample(transect_pool_nontreat, size = nrow(grid)/2, replace = TRUE)
pool_treat <- sample(transect_pool_treat, size = nrow(grid)/2, replace = TRUE)

library(merTools)

gridT <- cbind(grid, transect = c(pool_treat, pool_nontreat))

preds <- predictInterval(m3, newdata = gridT, type = "probability") # level = 0.8 by default
preds <- cbind(gridT, preds)

DamPercent <- preds %>% group_by(treatment) %>% 
  summarise(mean = mean(fit),
            low_CI = mean(lwr),
            high_CI = mean(upr))
DamPercent
```
# Number of acres of damage

```{r}
# All areas are in acres
# Treatment area:
areaT <- 199.86 # 80.88 ha
waterT <- 35.3274554562 # 14.296514 ha
roadT <- 1.716842761097904 #  6,947.816154 ha

areaNT <- 200.73 # 81.23 ha
waterNT <- 34.4680698895 # 13.948733 ha
roadNT <- 2.51733741370503 # 10,187.30308 m2

areaWT <- areaT - waterT - roadT
areaWNT <- areaNT - waterNT - roadNT

(damageT <- areaWT*DamPercent[DamPercent$treatment == "treatment",c("low_CI", "mean", "high_CI")])
(damageNT <- areaWNT*DamPercent[DamPercent$treatment == "non-treatment",c("low_CI", "mean", "high_CI")])

```



## B. BACI design

```{r}
data_BACI <- dataAll
data_BACI <- dataW

# For the BACI analysis, we only keep damage which is fresh or recent
data_BACI$Damage_Coverage[data_BACI$Age == 2] <- 0
```


```{r}
BACI_quadrats <- data_BACI %>% 
  group_by(Transect, Transect_Length_in_m, Before_After_Control, Observer) %>% 
  # Aggregate to the left and right (Width)
  summarise(successes = sum(Damage_Coverage), water = sum(Water_Coverage), total = n(), 
             dist_Fireline_m = mean(NEAR_DIST_Fireline_meters), 
             dist_AG_m = mean(NEAR_DIST_2021_NLCD_AG_meters), 
             NDWI = mean(NDWI), 
             dist_Burned_Unit_m = mean(NEAR_DIST_Burned_Unit_meters)) %>% 
  # Aggregateby series of 4 m (Length)
  group_by(Transect, Before_After_Control) %>% 
  mutate(quadrat = (row_number() - 1) %/% 4) %>%     
  group_by(Transect, Before_After_Control, quadrat) %>% 
  summarise(successes = sum(successes), water = sum(water), total = sum(total), Transect_Length_in_m = min(Transect_Length_in_m), 
             dist_Fireline_m = mean(dist_Fireline_m), 
             dist_AG_m = mean(dist_AG_m), 
             NDWI = mean(NDWI), 
             dist_Burned_Unit_m = mean(dist_Burned_Unit_m)) %>% 
  # Keep only complete quadrats (4m X 4m)
  filter(total == 32) %>% 
  # Keep only quadrats that are not covered 100% in water
  filter(water != 32) %>% 
  mutate(cells = total - water, failures = cells - successes) %>% 
  select(-total, -water, -cells)

# Width invariant data
Transect_data <- data_BACI[data_BACI$Transect_Width_in_m == 1 & data_BACI$Transect_Left_Right == "Right", c("Transect", "Before_After_Control", "Transect_Length_in_m", "Observer", "Treatment_Non_treatment", "Wetland_ATTRIBUTE", "WETLAND_TY", "EC_TYPE")]

BACI_quadrats <- left_join(BACI_quadrats, Transect_data, by = c("Transect", "Before_After_Control", "Transect_Length_in_m"))
# Create dummy variables for pre/post and treatment/non-treatment
BACI_quadrats$Post <- ifelse(BACI_quadrats$Before_After_Control == 'After', 1, 0)
BACI_quadrats$Treatment <- ifelse(BACI_quadrats$Treatment_Non_treatment == 'Treatment', 1, 0)


```

```{r}
library(lme4)
library(modelsummary) # Table functions
library(emmeans)

# Rescale distance to fire in km
BACI_quadrats$dist_fire_km <- BACI_quadrats$NEAR_DIST_Burned_Unit_meters/1000

yBACI <- cbind(BACI_quadrats$successes, BACI_quadrats$failures)

# Option 1: no transect random effect, no distance to fire
modelBACI <- glm(yBACI ~ Post + Treatment + Post*Treatment, family = binomial, data = BACI_quadrats)
coef(modelBACI, digits = 5)
exp(coef(modelBACI))
# (Intercept)           Post      Treatment Post:Treatment 
#  0.01189622     0.21220516     0.97903910     0.25744347 
# prob = odds / (1 + odds)
exp(confint(modelBACI))
# (Intercept)    0.01119624 0.01262512
# Post           0.18454333 0.24306235
# Treatment      0.88737148 1.07934006
# Post:Treatment 0.18526940 0.35115574
emmeans(modelBACI, ~ Post+Treatment, type='response')  # probabilities for each p/treatment

# Option 2: no transect random effect,  distance to fire
modelBACI <- glm(yBACI ~ Post + Treatment + I(dist_fire_km^2) + dist_fire_km + Post*Treatment, family = binomial, data = BACI_quadrats)
coef(modelBACI, digits = 5)
exp(coef(modelBACI))
# (Intercept)              Post         Treatment I(dist_fire_km^2)      dist_fire_km    Post:Treatment 
#   0.0140117         0.1820976         1.2365241         1.0299188         0.8035324         0.2535229 
# prob = odds / (1 + odds)
exp(confint(modelBACI))
# (Intercept)       0.01280509 0.01529942
# Post              0.15404323 0.21403968
# Treatment         1.08160387 1.41496266
# I(dist_fire_km^2) 1.01802332 1.04201109
# dist_fire_km      0.73962376 0.87251685
# Post:Treatment    0.18039922 0.35032273
emmeans(modelBACI, ~ Post+Treatment, type='response')  # probabilities for each p/treatment

# Option 3: transect random effect
modelBACI <- glmer(yBACI ~ Post + Treatment + Post*Treatment + (1|Transect), family = binomial, data = BACI_quadrats)
summary(modelBACI)
modelBACI@beta
exp(modelBACI@beta)
exp(confint(modelBACI))
#                    2.5 %    97.5 %
#.sig01         2.09848765 6.1615071
#(Intercept)    0.01399366 0.1152396
#Post           0.96944668 1.0644888
#Treatment      0.14153054 2.7927431
#Post:Treatment 0.75694484 0.8722369

emmeans(modelBACI, ~ Post+Treatment, type='response')  # probabilities for each p/treatment
tab_model(modelBACI)

# Option 3: transect random effect
modelBACI <- glmer(yBACI ~ Post + Treatment + dist_fire_km + Post*Treatment + (1|Transect), family = binomial, data = BACI_quadrats)
summary(modelBACI)
modelBACI@beta
exp(modelBACI@beta)
exp(confint(modelBACI))
#                    2.5 %    97.5 %
#.sig01         2.09848765 6.1615071
#(Intercept)    0.01399366 0.1152396
#Post           0.96944668 1.0644888
#Treatment      0.14153054 2.7927431
#Post:Treatment 0.75694484 0.8722369

emmeans(modelBACI, ~ Post+Treatment, type='response')  # probabilities for each p/treatment
library(sjPlot)
tab_model(modelBACI)

```

## (Intercept)     sexGirl  foodBreast   foodSuppl 
##   0.1993479   0.7315770   0.5120696   0.8415226
The interpretation here is that odds of respiratory infection for females is 73.1% than that of a similarly feed male child and I might say that being female reduces the odds of respiratory illness by 27 %  compared to male babies.
Chapter 11
https://bookdown.org/dereksonderegger/571/11-binomial-regression.html#interpreting-model-coefficients
